---
title: "Exploring Central Limit Theorem through Exponential Distribution Simulation"
author: "Han-Yu Hsieh"
date: "11/07/2020"
output:
  pdf_document: default
  html_document: default
---
### Overview
I would like to investigate the exponential distribution in R and compare it with the Central Limit Theorem. I set lambda as 0.2 for all of these simulations, which the theoretical mean is 5, and the theoretical variance is 0.625, and I generate a distribution of 1000 average of 40 exponentials, which I get a sample mean as 5.0059 and a sample variance as 0.602704. In addition, from the last table, we can see that the larger sample size, the distribution is more approximately normal because its sample mean and sample variance gets very very close to the true mean and variance. Therefore, the result of this simulation confirms the Central Limit Theorem.

### Simulations
I generated a distribution of 1000 average of 40 random exponentials with rexp function, and applied set.seed function to repeat this simulation.
```{r results='hide'}
library(ggplot2)
set.seed(20200709)
lambda=0.2;mean=1/lambda;sd=1/lambda
mns = NULL
for (i in 1:1000) mns = c(mns, mean(rexp(40,lambda)))
ep_data <- data.frame(mns, size = 40)
```
### Sample Mean versus Theoretic Mean
From the simulation that I created above, I would like to compare the sample mean and theoretic mean by plotting the simulation as a histogram by using ggplot2.
```{r}
exp_plot <- ggplot(ep_data, aes(x=mns, fill = size)) 
figure1 <- exp_plot+theme_bw()+
    geom_histogram(alpha=0.5, binwidth = 0.3, col="blue")+
    geom_vline(aes(xintercept = mean(mns), color = "red"))+
    geom_text(aes(x=mean(mns), label=mean(mns),y=50), angle=90)+
    theme(legend.position = "none")+
    xlab("Average of the distribution")+
    ylab("Frequence")+
    ggtitle("Figure 1: Distribution of the Average of 40 Random Exponentials(n=1000)")
figure1
```

```{r}
theoreticl_mean= 1/0.2
sample_mean_1= mean(ep_data$mns); sample_mean_1
```
From the mathematical equation, the mean of an exponential distribution is 1/lambda. Therefore, the theoretical mean would be 1/0.2, which is 5. As we can see from Figure 1, the red line indicates the sample mean, 5.0059, and it is very closed to the theoretical mean.
### Sample Variance versus Theoretic Variance
```{r}
theoreticl_variance=((1/0.2)/sqrt(40))^2
sample_variance_1=var(ep_data$mns); sample_variance_1
```
From the mathematical equation, the variance of a sample exponential distribution is (sd/sqrt(n))^2. Therefore, the theoretical variance is 0.625, where the calculated sample variance is 0.602704. The difference between the theoretical variance and sample variance largely depends on the sample size. When the sample size is larger, the difference between these two would be smaller. I demonstrated this concept in the next paragraph. 

### A large collection of sampling random exponentials
I generated a second distribution of 40 random exponentials by using the same method and the same parameter(lambda = 0.2). This time, I only increased the sample size to 10,000. 
```{r}
set.seed(20200710)
lambda=0.2;mean=1/lambda;sd=1/lambda
mns_2 = NULL
for (i in 1:10000) mns_2 = c(mns_2, mean(rexp(40,lambda)))
ep_data_2 <- data.frame(mns_2, size = 40)
exp_plot_2 <- ggplot(ep_data_2, aes(x=mns_2, fill = size)) 
figure2 <- exp_plot_2+theme_bw()+
    geom_histogram(alpha=0.5, binwidth = 0.3, col="dark green")+
    geom_vline(aes(xintercept = mean(mns_2), color = "red"))+
    geom_text(aes(x=mean(mns_2), label=mean(mns_2),y=500), angle=90)+
    theme(legend.position = "none")+
    xlab("Average of the distribution")+
    ylab("Frequence")+
    ggtitle("Figure 2: Distribution of the Average of 40 Random Exponentials(n=10,000)")
figure2
```
```{r}
result <- matrix(c(5, 0.625, mean(ep_data$mns),var(ep_data$mns), mean(ep_data_2$mns_2), var(ep_data_2$mns_2)),ncol = 3)
colnames(result) <- c("Theoretical", "n=1000", "n=10,000")
rownames(result) <- c("Mean", "Variance")
result <- as.table(result)
result
```

